# **GSE72540 Analysis**

## **WGCNA w/ all DEG**

<!-- Purge de l'environnement --->

```{r}
rm(list=ls())
```

<!-- Chargement et téléchargement de la biblio -->

```{r}
#BiocManager::install("WGCNA")
#devtools::install_github("kevinblighe/CorLevelPlot")
library(DESeq2)
library(GEOquery)
library(tidyverse)
library(CorLevelPlot)
library(gridExtra)
library(bbcViz)
library(WGCNA)
library(DT)
```

<!-- Chargement du eSet -->

```{r}
eset_file <- "GSE72540_eset.rds"
data <- read_rds(sprintf("/home/DA82_SUR/Internship_data/CSU/eSet/%s",eset_file))
```

<!-- Chargement du/des DEG (expression) -->

```{r}
DEG_exprs <- "DEG_expression_GSE72540.csv"
exprs_data <- read.csv(sprintf("/home/DA82_SUR/Internship_data/CSU/WGCNA/forWGCNA/%s",DEG_exprs))
rownames(exprs_data) <- exprs_data$X
exprs_data$X <- NULL
```

<!-- Open functions -->

```{r}
save_data <- T
source("~/Internship_data/CSU/wgcna_functions.R")
```

### **Visualization of input data**

:::info
Voici un extrait de la matrice d'expression sur laquelle nous lançons WGCNA (8 premières lignes et les 5 premières colonnes). Elle sera transposée au préalable pour pouvoir être utilisée dans l'algorithme.
:::
  
```{r}
# Convertir la matrice d'expression en data.frame pour une utilisation avec DT
expression_df <- as.data.frame(exprs_data)

# Use DT::datatable() to create an interactive table
createTable(expression_df[1:8,1:5])
```

<!-- Transposition de la matrice pour avoir les gènes en variables -->
  
:::info
Voici un extrait de la matrice transposée (5 premières lignes et 8 premières colonnes).
:::

```{r}
input_mat <- t(exprs_data)
input_mat <- as.data.frame(input_mat)
createTable(input_mat[1:5,1:8])         # Look at first 5 rows and 10 columns
```

<!-- Récupération des données de maladies des patients -->
  
```{r}
sample_status <-pData(data)[c("SamplePathology","DiseaseState","SampleID")]
sample_status$DiseaseState <- ifelse(sample_status$SamplePathology == "non-lesional" | sample_status$SamplePathology == "normal", "HC", sample_status$DiseaseState)
sample_status$DiseaseState <- ifelse(sample_status$DiseaseState != "HC", "CSU", sample_status$DiseaseState)
sample_status$SamplePathology <- NULL

```

<!-- Récupération des données de FC, p-value, ... -->
  
```{r}
DEA_data <- read.csv("/home/DA82_SUR/Internship_data/CSU/WGCNA/forWGCNA/DEG_GSE72540.csv")
DEA_FC <- DEA_data[c("Genes","FC")]
```

### **Pick threshold**{.tabset}

```{r include=FALSE}
allowWGCNAThreads()   # allow multi-threading (optional)
```

#### **Table of powers**{.tabset}

##### Unsigned
```{r,include=FALSE}
#Set network type
netwk_type_unsigned = "unsigned"
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to = 50, by = 2))
sft_data_unsigned <- pickThresholdAndPlot(netwk_type_unsigned,0)
```

```{r}
createTable(sft_data_unsigned)
```

##### Signed
```{r,include=FALSE}
#Set network type
netwk_type_signed = "signed"
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to = 50, by = 2))
sft_data_signed <- pickThresholdAndPlot(netwk_type_signed,0)
```

```{r}
createTable(sft_data_signed)
```

##### Signed hybrid
```{r,include=FALSE}
#Set network type
netwk_type_shybrid= "signed hybrid"
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to = 50, by = 2))
sft_data_shybrid <- pickThresholdAndPlot(netwk_type_shybrid,0)
```

```{r}
createTable(sft_data_shybrid)
```

#### **Sclale Independence and Mean Connectivity Plots**{.tabset}

##### Unsigned
```{r}
PlotsAndArrange(sft_data_unsigned)
```

##### Signed

```{r}
PlotsAndArrange(sft_data_signed)
```

##### Signed hybrid

```{r}
PlotsAndArrange(sft_data_shybrid)
```

```{r,echo=FALSE}
# Network Construction and Module Detection
picked_power_unsigned = determineOptimalPower(sft_data_unsigned,0.8)
netwk_unsigned <- constructNetworkAndDetectModules(netwk_type_unsigned,picked_power_unsigned,0)
```

```{r,echo=FALSE}
# Network Construction and Module Detection
picked_power_signed = determineOptimalPower(sft_data_signed,0.8)
netwk_signed <- constructNetworkAndDetectModules(netwk_type_signed,picked_power_signed,0)
```

```{r,echo=FALSE}
# Network Construction and Module Detection
picked_power_shybrid = determineOptimalPower(sft_data_shybrid,0.8)
netwk_shybrid <- constructNetworkAndDetectModules(netwk_type_shybrid,picked_power_shybrid,0)
```

### **Module Visualization** {.tabset}

#### Unsigned
:::info
Let's take a look at the modules, there
:::
```{r, fig.width=8, fig.height=4}
visualizeModules(netwk_unsigned)
```

#### Signed
:::info
Let's take a look at the modules, there
:::
```{r, fig.width=8, fig.height=4}
visualizeModules(netwk_signed)
```

#### Signed hybrid
:::info
Let's take a look at the modules, there
:::
```{r, fig.width=8, fig.height=4}
visualizeModules(netwk_shybrid)
```

### **Relate Module (cluster) Assignments to Sample** {.tabset}

#### **List of Module associate to Fold Change** {.tabset}

##### Unsigned

```{r}
#We can pull out the list of modules
module_df_unsigned <- relateModulesToTraits(netwk_unsigned)
```

:::info
Nous pouvons vérifier quels gènes sont présents dans quels modules et le FC associé au gène.
:::

```{r}
createTable(mergeFCWithModuleData(module_df_unsigned,DEA_FC))
```

##### Signed

```{r}
#We can pull out the list of modules
module_df_signed <- relateModulesToTraits(netwk_signed)
```

:::info
Nous pouvons vérifier quels gènes sont présents dans quels modules et le FC associé au gène.
:::

```{r}
createTable(mergeFCWithModuleData(module_df_signed,DEA_FC))
```

##### Signed hybrid

```{r}
#We can pull out the list of modules
module_df_shybrid <- relateModulesToTraits(netwk_shybrid)
```

:::info
Nous pouvons vérifier quels gènes sont présents dans quels modules et le FC associé au gène.
:::

```{r}
createTable(mergeFCWithModuleData(module_df_shybrid,DEA_FC))
```

#### **Summarize number of modules and genes** {.tabset}

##### Unsigned
:::info
Verifions combien de modules et combien de gènes on compte.
:::
```{r}
createTable(summarizeModuleData(module_df_unsigned))
#moduleSummary
#length(moduleSummary[,1])

```

##### Signed
:::info
Verifions combien de modules et combien de gènes on compte.
:::
```{r}
createTable(summarizeModuleData(module_df_signed))
#moduleSummary
#length(moduleSummary[,1])

```

##### Signed hybrid
:::info
Verifions combien de modules et combien de gènes on compte.
:::
```{r}
createTable(summarizeModuleData(module_df_shybrid))
#moduleSummary
#length(moduleSummary[,1])

```

#### **Heatmap of module and sample relationships** {.tabset}

##### Unsigned
:::info
We have written out a tab delimited file listing the genes and their modules. However, we need to figure out which modules are associated with each trait/treatment group. WGCNA will calcuate an Eigangene (hypothetical central gene) for each module, so it easier to determine if modules are associated with different treatments.
:::
```{r, fig.width=7, fig.height=8}
plotModuleTraitRelationships(input_mat,netwk_unsigned,sample_status,x = "treatment",y = "SampleID")$plot
```

```{r}
#Vérifier la liste des gènes par modules
list_of_modules_unsigned <- listOfModules(module_df_unsigned)
```

##### Signed
:::info
We have written out a tab delimited file listing the genes and their modules. However, we need to figure out which modules are associated with each trait/treatment group. WGCNA will calcuate an Eigangene (hypothetical central gene) for each module, so it easier to determine if modules are associated with different treatments.
:::
```{r, fig.width=7, fig.height=8}
plotModuleTraitRelationships(input_mat,netwk_signed,sample_status,x = "treatment",y = "SampleID")$plot
```

```{r,include=F}
# if(save_data){
#   write_rds(plotModuleTraitRelationships(input_mat,netwk_signed,sample_status,x = "treatment",y = "SampleID")$MEs,"~/Internship_data/CSU/Correlation/WGCNA_GSE72540_data_corr.rds")
# }
```

```{r}
#Vérifier la liste des gènes par modules
list_of_modules_signed <- listOfModules(module_df_signed)
```

##### Signed hybrid
:::info
We have written out a tab delimited file listing the genes and their modules. However, we need to figure out which modules are associated with each trait/treatment group. WGCNA will calcuate an Eigangene (hypothetical central gene) for each module, so it easier to determine if modules are associated with different treatments.
:::
```{r, fig.width=7, fig.height=8}
plotModuleTraitRelationships(input_mat,netwk_shybrid,sample_status,x = "treatment",y = "SampleID")$plot
```

```{r}
#Vérifier la liste des gènes par modules
list_of_modules_shybrid <- listOfModules(module_df_shybrid)
```

### **See modules of our genes of interest** {.tabset}

#### Unsigned

```{r, fig.width=20, fig.height=9}
# Examine Expression Profiles

#We'll pick out a few modules of interest, and plot their expression profiles 
#To do the choice we will pick modules in which our genes of interest are 

genes_of_interest <- c("TNFRSF4","CD40","CXCR3","CXCR4","BCL2A1","IL21R","TIGIT")

# Utiliser la fonction et stocker le résultat dans un data frame
genes_in_modules_df_unsigned <- group_genes_by_module(genes_of_interest, list_of_modules_unsigned)

# Afficher le résultat
createTable(genes_in_modules_df_unsigned)
```

#### Signed

```{r, fig.width=20, fig.height=9}
# Examine Expression Profiles

#We'll pick out a few modules of interest, and plot their expression profiles 
#To do the choice we will pick modules in which our genes of interest are 

genes_of_interest <- c("TNFRSF4","CD40","CXCR3","CXCR4","BCL2A1","IL21R","TIGIT")

# Utiliser la fonction et stocker le résultat dans un data frame
genes_in_modules_df_signed <- group_genes_by_module(genes_of_interest, list_of_modules_signed)

# Afficher le résultat
createTable(genes_in_modules_df_signed)
```

#### Signed hybrid

```{r, fig.width=20, fig.height=9}
# Examine Expression Profiles

#We'll pick out a few modules of interest, and plot their expression profiles 
#To do the choice we will pick modules in which our genes of interest are 

genes_of_interest <- c("TNFRSF4","CD40","CXCR3","CXCR4","BCL2A1","IL21R","TIGIT")

# Utiliser la fonction et stocker le résultat dans un data frame
genes_in_modules_df_shybrid <- group_genes_by_module(genes_of_interest, list_of_modules_shybrid)

# Afficher le résultat
createTable(genes_in_modules_df_shybrid)
```

### **Examine profiles of modules**{.tabset}

#### Unsigned

```{r}
# pick out a few modules of interest here
modules_of_interest_unsigned <- c("turquoise", "blue","grey","brown",genes_in_modules_df_unsigned$module)
modules_of_interest_unsigned <- unique(modules_of_interest_unsigned)

# Pull out list of genes in that module
submod_unsigned = module_df_unsigned %>%
  subset(colors %in% modules_of_interest_unsigned)

row.names(module_df_unsigned) = module_df_unsigned$gene_id

#Examine profiles
examineExpressionProfiles(submod_unsigned,module_df_unsigned)
```

#### Signed

```{r}
# pick out a few modules of interest here
modules_of_interest_signed <- c("turquoise", "blue","grey","brown",genes_in_modules_df_signed$module)
modules_of_interest_signed <- unique(modules_of_interest_signed)

# Pull out list of genes in that module
submod_signed= module_df_signed %>%
  subset(colors %in% modules_of_interest_signed)

row.names(module_df_signed) = module_df_signed$gene_id

#Examine profiles
examineExpressionProfiles(submod_signed,module_df_signed)

```

#### Signed hybrid

```{r}
# pick out a few modules of interest here
modules_of_interest_shybrid <- c("turquoise", "blue","grey","brown",genes_in_modules_df_shybrid$module)
modules_of_interest_shybrid <- unique(modules_of_interest_shybrid)

# Pull out list of genes in that module
submod_shybrid = module_df_shybrid %>%
  subset(colors %in% modules_of_interest_shybrid)

row.names(module_df_shybrid) = module_df_shybrid$gene_id

#Examine profiles
examineExpressionProfiles(submod_shybrid,module_df_shybrid)

```

### **Find hub genes for each modules**{.tabset}

#### Unsigned
```{r,echo=F}
top_hub_genes_unsigned <- chooseTopHubGenes(list_of_modules_unsigned, input_mat, picked_power_unsigned, netwk_type_unsigned,"grey")

# Afficher les gènes hub dans une table
hub_genes_df_unsigned <- data.frame(Module = character(), HubGene = character(), stringsAsFactors = FALSE)
for (module in names(top_hub_genes_unsigned)) {
  hub_genes <- top_hub_genes_unsigned[[module]]
  if (length(hub_genes) > 0) {
    for (gene in hub_genes) {
      hub_genes_df_unsigned <- rbind(hub_genes_df_unsigned, data.frame(Module = module, HubGene = gene))
    }
  }
}
createTable(hub_genes_df_unsigned)
```

#### Signed
```{r,echo=F}
top_hub_genes_signed <- chooseTopHubGenes(list_of_modules_signed, input_mat, picked_power_signed, netwk_type_signed,"grey")

# Afficher les gènes hub dans une table
hub_genes_df_signed <- data.frame(Module = character(), HubGene = character(), stringsAsFactors = FALSE)
for (module in names(top_hub_genes_signed)) {
  hub_genes <- top_hub_genes_signed[[module]]
  if (length(hub_genes) > 0) {
    for (gene in hub_genes) {
      hub_genes_df_signed <- rbind(hub_genes_df_signed, data.frame(Module = module, HubGene = gene))
    }
  }
}
createTable(hub_genes_df_signed)
```

#### Signed hybrid
```{r,echo=F}
top_hub_genes_shybrid <- chooseTopHubGenes(list_of_modules_shybrid, input_mat, picked_power_shybrid, netwk_type_shybrid,"grey")

# Afficher les gènes hub dans une table
hub_genes_df_shybrid <- data.frame(Module = character(), HubGene = character(), stringsAsFactors = FALSE)
for (module in names(top_hub_genes_shybrid)) {
  hub_genes <- top_hub_genes_shybrid[[module]]
  if (length(hub_genes) > 0) {
    for (gene in hub_genes) {
      hub_genes_df_shybrid <- rbind(hub_genes_df_shybrid, data.frame(Module = module, HubGene = gene))
    }
  }
}
createTable(hub_genes_df_shybrid)
```


### **Display edge list with correlation values**{.tabset}

#### Unsigned

```{r,include=FALSE}
genes_of_interest_unsigned = module_df_unsigned %>%
  subset(colors %in% modules_of_interest_unsigned)

expr_of_interest_unsigned = exprs_data[genes_of_interest_unsigned$gene_id,]


# Recalculate TOM for modules of interest 
list_unsigned <- calculateTOMandEdgeList(module_df_unsigned,picked_power_unsigned,modules_of_interest_unsigned)

edge_list_unsigned <- list_unsigned$edge_list
TOM_unsigned <- as.data.frame.matrix(list_unsigned$TOM)

writeEdgeList(eset_file,netwk_type_unsigned,edge_list_unsigned)



# Boucle pour générer et enregistrer les edge lists pour chaque module
for (module in unique(module_df_unsigned$colors)) {
  # Filtrer l'edge list pour un module spécifique
  edge_list_filtered_unsigned <- edge_list_unsigned %>%
    dplyr::filter(module1 == module & module2 == module)
  
  # Sauvegarder l'edge list filtrée dans un fichier
  writeEdgeListForModule(edge_list_filtered_unsigned, module, netwk_type_unsigned, eset_file)
}
```
:::info
Voici un extrait de la liste des corrélations entre gènes
:::
```{r}
createTable(edge_list_unsigned[1:2000,])
```

#### Signed

```{r,include=FALSE}
genes_of_interest_signed = module_df_signed %>%
  subset(colors %in% modules_of_interest_signed)

expr_of_interest_signed = exprs_data[genes_of_interest_signed$gene_id,]

# Recalculate TOM for modules of interest 
list_signed <- calculateTOMandEdgeList(module_df_signed,picked_power_signed,modules_of_interest_signed)

edge_list_signed <- list_signed$edge_list
TOM_signed <- as.data.frame(list_signed$TOM)

writeEdgeList(eset_file,netwk_type_signed,edge_list_signed)


# Boucle pour générer et enregistrer les edge lists pour chaque module
for (module in unique(module_df_signed$colors)) {
  # Filtrer l'edge list pour un module spécifique
  edge_list_filtered_signed <- edge_list_signed %>%
    dplyr::filter(module1 == module & module2 == module)
  
  # Sauvegarder l'edge list filtrée dans un fichier
  writeEdgeListForModule(edge_list_filtered_signed, module, netwk_type_signed, eset_file)
}
```

:::info
Voici un extrait de la liste des corrélations entre gènes
:::
```{r}
createTable(edge_list_signed[1:2000,])
```


#### Signed hybrid

```{r,include=FALSE}
genes_of_interest_shybrid = module_df_shybrid %>%
  subset(colors %in% modules_of_interest_shybrid)

expr_of_interest_shybrid = exprs_data[genes_of_interest_shybrid$gene_id,]

# Recalculate TOM for modules of interest 
list_shybrid <- calculateTOMandEdgeList(module_df_shybrid,picked_power_shybrid,modules_of_interest_shybrid)

edge_list_shybrid <- list_shybrid$edge_list
TOM_shybrid <- as.data.frame(list_shybrid$TOM)

writeEdgeList(eset_file,netwk_type_shybrid,edge_list_shybrid)


# Boucle pour générer et enregistrer les edge lists pour chaque module
for (module in unique(module_df_shybrid$colors)) {
  # Filtrer l'edge list pour un module spécifique
  edge_list_filtered_shybrid <- edge_list_shybrid %>%
    dplyr::filter(module1 == module & module2 == module)
  
  # Sauvegarder l'edge list filtrée dans un fichier
  writeEdgeListForModule(edge_list_filtered_shybrid, module, netwk_type_shybrid, eset_file)
}

```

:::info
Voici un extrait de la liste des corrélations entre gènes
:::
```{r}
createTable(edge_list_shybrid[1:2000,])
```

```{r,include=F}
# Save module for Lasso Regression
turquoise_module_GSE72540 <- rownames(module_df_signed[which(module_df_signed$colors=="turquoise"),])
blue_module_GSE72540 <- rownames(module_df_signed[which(module_df_signed$colors=="blue"),])
grey_module_GSE72540 <- rownames(module_df_signed[which(module_df_signed$colors=="grey"),])
brown_module_GSE72540 <- rownames(module_df_signed[which(module_df_signed$colors=="brown"),])


save(turquoise_module_GSE72540,blue_module_GSE72540,file="modules_list_GSE72540.RData")
```
